---
title: "Bureaucracy proj (demo)"
author: "Anastasiia Karpenko"
date: "18 09 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Dealing with government bureaucracy is not the most pleasant thing in the world. In this demo project I am looking at the association between the time the bureaucracy consumes and different economic indicators. In particular, to measure the bureaucracy I took the number of days and procedures one has to spend to register a property from the Doing Business Index.

I collected data for 145 countries for four years (2008, 2011, 2014, and 2017) from open resources like World Bank, Doing Business Index, International Labor Organization, etc.

```{r libraries, include=FALSE}
# load libraries

loadmanylibs <- c('ggplot2','corrplot','gridExtra', 'dplyr', 'car', 'lmtest', 'statmod',
                  'ggalt', 'dplyr', 'readxl', 'plm', 'data.table', 'sandwich',
                  'writexl', 'lares', 'ggpubr', 'stargazer', 'fastDummies', 'Hmisc')
#install.packages(loadmanylibs)
lapply(loadmanylibs, require, character.only = TRUE)
```

```{r data preparation, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How does the bureaucracy associate with different economic indicators?

For example, a higher number of procedures one has to complete to register a property is associated with higher (lower) level of (absence) corruption, lower income and less democratic regimes.

```{r corrplot, echo=FALSE, fig.height=9, fig.width=9}
load("C:/Users/пк/Desktop/bureaucracy/processed_data/cleandata.RData")
num_cols_pr <- unlist(lapply(present, is.numeric))
corrplot_pr <- present[,num_cols_pr] # correlations of all numeric variables
cor_numvar_pr <- cor(corrplot_pr, use="pairwise.complete.obs")
colnames(cor_numvar_pr) <- 
  c("Government spending efficiency", "Government debt % of gdp", "Budget balance % of gdp", "Budget transparency", "Government effectiveness",
    "Democracy", "GDP growth", "GDP per capita", "Unemployment", "Procedures to register a property",
    "Days to register a property", "Absence of corruption")
rownames(cor_numvar_pr) <-colnames(cor_numvar_pr)

corrplot(cor_numvar_pr, type="full", tl.col="black", method="color")
```

##  Does more bureaucracy mean better spending of budget resources?

The figure below shows that more bureaucratic checks do not help much the government to distribute its finances more efficiently.

```{r linear-plot, echo=FALSE, fig.height=7, fig.width=9}
grid.arrange(
  
  ggplot(present,aes(x = days, y = Efficiency)) + 
    geom_point() + 
    labs(x = 'Number of days to register a property') +
    labs(y = 'Government spending efficiency') +
    geom_smooth(method = "loess", se=TRUE) +
    theme_classic() + theme(axis.text = element_text(size = 10)) +
    stat_regline_equation(label.y = 6, label.x = 200, aes(label = ..eq.label..)) +
    stat_regline_equation(label.y = 5, label.x = 200, aes(label = ..rr.label..)),
  
  ggplot(present,aes(x = procedures, y = Efficiency)) + 
    geom_point() + 
    labs(x = 'Number of procedures to register a property') +
    labs(y = 'Government spending efficiency') +
    geom_smooth(method = "loess", se=TRUE) +
    theme_classic() + theme(axis.text = element_text(size = 10)) +
    stat_regline_equation(label.y = 6, label.x = 8, aes(label = ..eq.label..)) +
    stat_regline_equation(label.y = 5, label.x = 8, aes(label = ..rr.label..)),
  
  ncol = 1, nrow = 2)
```

## What is the government spending efficiency per income group?

## What is the share of government debt to GDP per income group?

On average, countries that fell in a high-income group have higher efficiency of government spending as well as higher share of government debt to GDP.

```{r grouped-plot, echo=FALSE, fig.height=7, fig.width=9}
load("C:/Users/пк/Desktop/bureaucracy/processed_data/yrav_byincgr.RData")

# in Efficiency of government spending
# in % of government debt to GDP by income_group
rename <- c(`High_income` = "High-income",`Low_income` = "Low-income",
            `Lower_middle_income`="Lower middle-income",
            `Upper_middle_income`="Upper middle-income")
grid.arrange(
  ggplot(data=mean_yr_inc_pr, aes(x=as.factor(Year), y=Efficiency, group=income_group))+
    geom_line() + geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = F) + labs(x='Year') + 
    labs(y='Efficiency of government spending') + theme_classic() +
    scale_y_continuous()+
    facet_wrap(~income_group, labeller = as_labeller(rename), ncol = 4),
  ggplot(data=mean_yr_inc_pr, aes(x=as.factor(Year), y=govdbtshgdp, group=income_group))+
    geom_line() + geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = F) + labs(x='Year') + 
    labs(y='% of government debt to GDP') + theme_classic() +
    scale_y_continuous()+
    facet_wrap(~income_group, ncol = 4, labeller = as_labeller(rename)), ncol = 1, nrow = 2)
```